---
title: "Describe"
author: "Dean Stewart"
date: "18/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Mortality Dataset

Summarising information about the French and swiss mortality  dataset.  

```{r, include= FALSE}
motor <- read_csv("1_data/freMTPL2freq.csv")
```

```{r}
glimpse(motor)

```

```{r}
summary(motor)
library(psych)
describe(motor)
```

```{r}
cor(motor[,c(0,1,2,3,5,6,7,8,11)])
motor$Area <- as.factor(motor$Area)
motor$VehBrand <- as.factor(motor$VehBrand)
motor$VehGas <- as.factor(motor$VehGas)
motor$Region <- as.factor(motor$Region)
str(motor)
summary(motor)
```

```{r}
prop.table(table(motor$VehBrand))
whatsthis <- data.frame(prop.table(table(motor$Area)))
whatsthis %>% filter(whatsthis$Var1=="F")
whatsthis[whatsthis$Var1=="F",]
```
## Week 4 Question 2
2i
```{r}
glimpse(motor)

#question talks of claim frequency assume this means ClainNb / Exposure

motor <- mutate(motor,ClaimFreq = ClaimNb/Exposure)

library(ggplot2)
motor %>%
  ggplot(aes(Exposure)) +
  geom_histogram(binwidth=1)

ggplot(motor,aes(Exposure)) +
  geom_histogram()

ggplot(motor,aes(Exposure)) +
  geom_boxplot()

ggplot(motor,aes(ClaimNb)) +
  geom_histogram()

#reading the marking guide tells us what these variables are
#cap esposure at 1 and cap numbers at 4 'because the larger ones ae probably data errors'

motor$Exposure <- pmin(motor$Exposure,1)
motor$ClaimNb <- pmin(motor$ClaimNb,1)
motor <- mutate(motor,ClaimFreq = ClaimNb/Exposure)

```

What do I see?  A low number a claims but a wide variation in exposure.

2ii
```{r}
overall_freq = sum(motor$ClaimNb) / sum(motor$Exposure)  
overall_freq

```
2iii
```{r}
glimpse(motor)
glimpse(motor[,c(2,3,5:8)])
#pairs(motor[,c(2,3,5:8,11,13)])  DON'T DO THIS!! TOO SLOW
library(corrplot)
corrplot(cor(motor[,c(2,3,5:8,11,13)]), use="pairwise.complete.obs")

#column plots
ggplot(motor,aes(Exposure, Area)) +
  geom_col()

ggplot(motor,aes(Exposure, VehPower)) +
  geom_col()

ggplot(motor,aes(Exposure, VehAge)) +
  geom_col()

ggplot(motor,aes(Exposure, DrivAge)) +
  geom_col()

ggplot(motor,aes(Exposure, BonusMalus)) +
  geom_col()

ggplot(motor,aes(Exposure, VehBrand)) +
  geom_col()

ggplot(motor,aes(Exposure, VehGas)) +
  geom_col()

ggplot(motor,aes(Exposure, Region)) +
  geom_col()
         
```

Bonus Malus is negatively correlated with driver age
Exposure is higher for older drivers and vehicles
high exposure areas are low bonus malus areas
no major correlations between claimFreq and anything else

Area F has a low exposure
large variation on brand and region

continuous variables don't plot - of couare.


```{r}
#scatter plots
ggplot(motor,aes(ClaimFreq, Area)) +
  geom_point()

ggplot(motor,aes(ClaimFreq, VehPower)) +
  geom_point()

ggplot(motor,aes(ClaimFreq, VehAge)) +
  geom_point()

ggplot(motor,aes(ClaimFreq, DrivAge)) +
  geom_point()

ggplot(motor,aes(ClaimFreq, BonusMalus)) +
  geom_point()

ggplot(motor,aes(ClaimFreq, VehBrand)) +
  geom_point()

ggplot(motor,aes(ClaimFreq, VehGas)) +
  geom_point()

ggplot(motor,aes(ClaimFreq, Region)) +
  geom_point()
         
```

There are a couple of regions with outlier claim frequency (probably regions withlow exposures)
Vehicle power and claim frequency seems negatively correlated
VEhicle age also seens negatively correlated.  

Are there confounders?

younger drivers have more claims - traditiona wisdom

those with more claims have less bonus - duh

2iv
```{r}
summary_density <- motor %>% mutate(logden = log(Density))

summary_density <- summary_density %>%
  mutate(logdengroup = case_when(logden<=1~1,
                    logden>1&logden<=2~2,
                    logden>2&logden<=3~3,
                    logden>3&logden<=4~4,
                    logden>4&logden<=5~5,
                    logden>5&logden<=6~6,
                    logden>6&logden<=7~7,
                    logden>7&logden<=8~8,
                    logden>8&logden<=9~9,
                    logden>9~10))
summary_density$logdengroup <- as.integer(summary_density$logdengroup)

logden_exposure <- ggplot(summary_density,aes(as_factor(logdengroup),Exposure)) +
  geom_boxplot()


summary_density %>% select(logdengroup,ClaimFreq) %>%
  group_by(logdengroup) %>%
  mutate(meanfreq = mean(ClaimFreq)) %>%
  ggplot(aes(as_factor(logdengroup),meanfreq)) +
  geom_point()

logden_freq <- summary_density %>% select(logdengroup,ClaimFreq) %>%
  group_by(logdengroup) %>%
  mutate(meanfreq = mean(ClaimFreq)) %>%
  ggplot(aes(logdengroup,meanfreq)) +
  geom_point()

library(gridExtra)
grid.arrange(logden_exposure,logden_freq)

```

